PhD Title \\ PhD Thesis

### 2024-06-07 Latex_files/Thesis/chapters/chapter 1/chapter1.tex
updating main and chapters using

\chapter{Introduction}
Crystallisation, put simply, is the formation and growth of a new structured phase within a disordered bulk phase. This has applications in a 
number of industries such as pharmaceuticals, food production, and 
electronics; where the extraction of dilute materials can help improve 
product quality while keeping production costs low. The industrialisation 
of crystallisation has allowed engineers to reliably and efficiently 
induce the crystal formation within a bulk liquid. However, while on a 
large scale crystallisation is seemingly an understood physical process, 
just a small amount of investigation into the literature reveals that at 
a micro level there is yet to be unifying theory that can accurately 
explain the process of crystallisation, more specifically nucleation.

Nucleation is the initial formation of the crystalline phase and can be 
broadly divided into two main processes. Firstly there is primary 
nucleation, in which the nucleus forms from the bulk liquid; and 
secondary nucleation, where the presence of already existing crystals 
promotes the formation and growth of new nuclei

\section{Nucleation Theories}
Nucleation is a common example of a binary phase separation, where a dilute phase is miscible in a bulk phase solute and solvent respectfully. While the solution remains supersaturated there is a chemical potential driving force for the solute to coalesce and separate from the solution as a crystal, the first formation of the crystal is referred to as the nucleus and understanding its formation of has been the focus of researchers for decades now. In particular there is a growing desire for a concrete theory to describe the process of nucleation.

\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the original theory was first formed from the works of Volmer and Weber, Becker and Doring, and Frenkel \cite{Frenkel1939, Volmer1926}. While initially it was more focused into describing droplet formation in condensing vapours it was extrapolated to describe crystallisation. The central premise is that in order for a nucleus to form within the bulk phase there must be some local fluctuation in the concentration of the solute. Now the nucleus will result in a local increase in potential ($\Phi$) but will also be subject to surface tension ($\gamma$) trying to dissolve the nucleus. Therefore the free energy change of a nucleus of size $r$ is given by:
\begin{align}
	\Delta G = \Phi r - \gamma r^{2/3}
\end{align}
From the equation its clear that there must be some critical size of nucleus above which the potential change is always positive. This critical size defines the free energy barrier that determines whether a nucleus will grow or dissolve within the solution. With this in mind the rate of which stable nuclei are formed can be approximated using the Arrhenius reaction rate equation:
\begin{align}
	J = A \ exp \left[-\frac{\Delta G^*}{k_BT} \right]
\end{align}
Where A is a prefactor that accounts for kinetic contributions. One of the key assumptions made in the original theory is that the nucleus' shape is always spherical, making calculation of the surface tension rather simple. Since the potential change is influenced by the supersaturation the nucleation rate should increase exponentially with supersaturation, and likewise with temperature (assuming the same thermodynamic supersaturation). In addition the theory implies that any collection of solute molecules below the critical size are unstable.

Conceptually this is a rather idealised theory as it captures all of the microscopic effects into only two terms and provides a rather simple view of nucleation as being a single step process. However, in practice there are a number of challenges to applying this to experimental measurements.

Where $A$ is a pre-factor that can be fine tuned to the exact demands of the system (typically involving Zeldovich Factor, $Z$) $\Delta G^*$ is the free energy barrier, this can be found by finding the point where the gradient goes to 0:
\begin{align}
	\Delta G = \gamma n^{2/3}-\phi n \\
	\frac{\delta G}{\delta n} = \frac{2}{3n^{1/3}}\gamma - \phi  = 0 \\
	n^* = \left( \frac{2\gamma}{3\phi} \right)^3 \\
	\Delta G^* = \gamma (n^*)^{2/3} - \phi n^*
\end{align}
So to find the critical size $n^*$  - and therefore the nucleation rate - we simply need to know the ratio between the interfacial free energy and the volume free energy. The CNT assumes that the crystal grows is purely spherical and that the interfacial free energy is near constant (aka that of a flat surface), this assumption is not very accurate for small crystal nucleus where the changing curvature means that the surface tension varies as a function of the crystal radius. This also assumes the surrounding fluid is at rest and there are no external forces on the nucleus that might distort the shape of said nucleus; though there has been theoretical work looking into how the CNT behaves under shear flow \cite{Debuysschere2023, Mura2016, Richard2015}

Furthermore, multiple experiments have already verified that crystals can form any number of structures and can even transition through multiple structures as it grows, meaning that the interfacial tension will constantly change as the particle grows. Furthermore, because we are assuming that the surface energy is constant the pre-exponential factor that is computed from experimental data, and the factor predicted by theory often disagree with one another to several degrees of magnitude.

There are several results that indicate that the CNT is not a good fit for nucleation rates:
\textcolor{red}{
	\begin{itemize}
		\item Discrepancy in vapor-liquid nucleation systems to the extent that CNT was off by a factor of more than 2000 [ ]. 
		\item In glass formation, nucleation kinetics are incredibly important to understand as the crystallization process is not desired when producing glass. We want to cool the molten solution so that the solid does not have time to reorganize itself into an ordered structure. To understand nucleation kinetics in glass formation we need to understand that as crystals form in the glass the surface energy differs from the stable phase significantly which us poorly predicted by CNT as it assumes that crystals will have a constant structure and interaction with the bulk material [ ]. 
		\item Metal alloys can demonstrate super-cooling while not experiencing nucleation, this is because in these liquids exhibit short range dynamics which means that the liquid droplet remains unordered as it is cooled. Because CNT assumes that the droplets must be of a similar ordering to that of the bulk crystal it often struggles to consider systems where the local bond order can fluctuate between phases. [ ]
	\end{itemize}
}

CNT is often regarded as a good description of the macro system, its obvious that for all crystallization systems there is an inherent energy barrier that dictates the nucleation rate. However, for such a diverse range of possible nucleation events the CNT crudely assumes all interactions between solutes as identical; this makes it difficult to model crystallisation events using CNT without significantly modifying the model for each specific case. Recent developments in studying nucleation events have opened up new possible 'pathways' for a nucleation event to go down, by far the most common is the idea of Two step nucleation.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.  It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards. This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state. Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation. The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.

\section{Optical Tweezers}
\subsection{Background}
Optical tweezing has been a field of applied optics ever since the 1970s when Ashkin \cite{Ashkin1970} first showed that focused light was capable of exerting 'radiation pressure'. The working principle was that a light source such as a laser could trap small objects within a 2D plane, as long as the light source had a symmetrical profile and whose intensity was concentrated around a central axis. Soon after, Ashkin showed that the introduction of a microscope objective would allow one to focus the light source to a diffraction limited point that would stably trap small objects within a confined volume \cite{Ashkin1980}. This allowed Ashkin and others to study biological material and would later be used to probe microscopic properties such as the formation of colloidal aggregates \cite{Yi2021} to the drag forces exerted by a pure vacuum \cite{Ahn2018, Monteiro2018}. Due to the predictable behaviour of light, optical tweezers have become essential for measuring and exerting precise forces on the magnitude of pico-newtons allowing one to probe the material properties of the smallest materials.

\subsection{Literature related to laser induced nucleation}
From as early as 1996 it has been known that laser irradiation is a viable method of inducing nucleation within a supersaturated solution \cite{Garetz1996}, the first reported case was notable as it used a 1.064 $\mu m$ laser meaning it was unlikely to be a photo-chemical reaction but rather a physical one. Future research has found nucleation can be induced by 1 of 3 routes involving direct laser induction: firstly, Non-photochemical laser induced nucleation (NPLIN) where the solution is irradiated with a pulsed laser \cite{Garetz1996, Garetz2002,Sun2006}, several papers have debated the exact mechanism that induces NPLIN \cite{Garetz2002, Knott2011}. Two suggested hypothesis are: an optical Kerr effect, where the solute molecules are aligned to lower the nucleation barrier \cite{Knott2011}; a dielectric polarisation effect, in which solute clusters are stabilised within the electric field which drastically increases the likelihood of nucleation\cite{Alexander2008}. Both hypothesise have their limitations and there has yet to be a single theory that explains NLPIN thoroughly. In either case the mean pulse intensity needs to be kept relatively low (on the order of $0.1-0.01 GW/cm^2$), as high intensity pulses lead to a completely different nucleation mechanism.

High intensity laser induced nucleation (HILIN), where the pulse intensity is on the order of several $PW/cm^2$ is far simpler a mechanism to explain in comparison to NPLIN. The production of nuclei can be wholly associated to a cavitation process within the target solution, where the laser focus results in thermo-cavitation and the subsequent pressure change leads to a nucleation event around the focus of the laser \textcolor{red}{[ , , ]}. There is still not a general consensus on how cavitation influences the local supersaturation \textcolor{red}{[ , ]}, nor is there a clear understanding of what properties of the crystal can be controlled \textcolor{red}{[ , ]}. It has been suggested that in theory any solution can undergo HILIN \textcolor{red}{[ ]}, proving such a theory requires a strong understanding of the phenomena both before and after cavitation occurs.

Lastly, there is trapping induced nucleation, this is where optical tweezers come into play, the optical trap has been shown to have different effects on supersaturated solutions depending on where it has been focused. When focusing on the cover slip, supersaturated solutions of glycine and $D_2O$ were shown to create a dense liquid droplet of glycine and water \textcolor{red}{[ , ]}, applying DLS analysis showed that the dense liquid region was populated by clusters that would consolidate together upon being focused by the optical trap \textcolor{red}{[ ]}. Molecular simulations of glycine solutions showed that these clusters are unstable when using pure glycine below the saturation point \textcolor{red}{[ ]} suggesting that the clusters are formed due to glycine reaction products. When the optical trap was moved from the cover slip to the air-solution interface nucleation would occur before a dense liquid region could form \textcolor{red}{[ ]}. Repeated experiments where the laser is focused on the air-solution interface have lead to a variety of different nucleation events. In some instances the nucleation occurs spontaneously after a short period of time \textcolor{red}{[ ]}; whereas allowing a solution to age results in the formation of amorphous precursors that when irradiated will nucleate immediately \textcolor{red}{[ ]}. The precursors are only seen when the solution is irradiated by an optical tweezer and the growth rate can be controlled somewhat by varying the laser power \textcolor{red}{[ ]}. Notably the only work has been done with simply irradiating the solution with a trapping potential, there has not been an attempt to introduce a trappable object into the solution, the trapping potential has been used to influence the growth of a crystal front \textcolor{red}{[ ]}.

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and angular momentum \textcolor{red}{[ ]}; more accurately the field is said to have both orbital and spin momentum. Though there is some debate on how to decompose the total momentum into these two components \textcolor{red}{[ ]}, for this thesis we do not need to calculate the exact quantities and will instead look at the broader effects of both components. The orbital momentum can be understood as the shape of the wavefront of the particular field in question, for simple Gaussian beams the wavefronts are uniform and equally spaced resulting in the typical radiation pressure that Ashkin and co demonstrated \textcolor{red}{[ ]}. However, higher order modes of a Gaussian beam (such as the Laguerre-Gaussian modes) have non-uniform wave fronts meaning the orbital momentum has both angular and linear components; depending on the relative size of the target particle one can induce rotation, or orbiting \textcolor{red}{[ ]}. Spin angular momentum (SAM) is attributed to the spin density of the field, early research has shown that the spin density is non-zero for any beam but while the SAM can easily be zero for homogeneous scatterers, this has sparked debate if SAM is even a physical quantity as it does not aid in the transport of energy directly \textcolor{red}{[ ]}. This paradox is resolved by representing the wave as an array of finite loops that all together cancel one-another out when the medium is homogeneous, when inhomogeneity is introduced (such as by being refracted by a birefringent medium) these circular loops no longer cancel out resulting in a non-zero SAM \textcolor{red}{[ ]}, depending on the material a number of rotational effects can occur.

For a perfectly homogeneous non-absorbing sphere the total SAM transferred from a circularly polarised beam is zero; if, however, the wave front of the beam is helical - for example using a Laguerre-Gaussian beam \textcolor{red}{[ ]} - one can control the rotational motion via transfer of OAM. Due to the structure of the LG beam the individual rays are not perpendicular with the propagation direction; as a trapped particle will experience a constant torque while being pulled towards the centre of the trap focus. This was demonstrated in \textcolor{red}{[ ]} where Copper Oxide micro particles rotated with a frequency of around 2 Hz due to the orbital AM from a LG beam; this was enhanced further by circularly polarising the trapping beam. Typically for a non-absorbing particle the polarization state of the beam has a negligible impact on the angular momentum transferred, this is not the case for particles with a high absorption efficiency as the combined handedness of the LG beam and polarisation vector can lead to an enhanced transfer of SAM. However, by far the best method for inducing rotation is using birefringent materials.

Birefringence is a material property often seen in crystalline materials, if the crystal lattice has a different structures when viewed at different orientations then light will be refracted differently depending on its polarisation. For circularly polarised light the inhomogeneity results in a high degree of SAM being transferred to the target object \textcolor{red}{[ ]}, this has been exploited to rotate microspheres as fast as 1000 Hz while suspended in a bulk medium \textcolor{red}{[ ]} as well as a means of measuring the local temperature and shear response of said medium \textcolor{red}{[ , ]}. Calculating the optical torque applied to a birefringent material is given via:

\begin{equation}
	\label{eq:opt_torque}
	\begin{aligned}
		\tau_{opt} =& -\frac{\epsilon}{2\omega_{laser}}E_0^2sin(kd(\Delta n))cos2\theta sin2\phi \\ &+  \frac{\epsilon}{2\omega_{laser}}E_0^2 (1-cos(kd(\Delta n))sin2\phi)
	\end{aligned}
\end{equation}

Where $\theta$ is the angle between the particle's orientation vector and the propagation direction of the EM field, and $\phi$ is the phase shift in the EM field.  The first term represents the 'orientational' torque which is due to the target particle being aligned along the EM field, when aligned $\theta=0$ meaning the entire term is negligible for particle's with a stable orientation. The second term is due purely to the polarisation of the optical trap, for circularly polarised light $\phi=\pi/4$ thus maximising the torque transferred to the target particle. Birefringence can also be induced if the target particle has an anisotropic shape, in particular if the particle shape is elongated along one major axis; the go to particle shape is a silica dimer (two spheres tangentially attached) due to silica's stability and strong adhesion. Using a silica dimer research groups have achieved a rotation frequency in the realm of several GHz \textcolor{red}{[ ]} in a vacuum. 
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

>>>>

\include{main}
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 3/chapter3.tex
updating main and chapters using

\setcounter{chapter}{2}
\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution. The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results. It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}. Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}. These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.

\section{Optical Tweezer Equipment}
In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals. The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser. Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture. To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications. In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification. Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging. By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations. While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement. The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane. Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm. To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.

\subsection{Position detection methods}
There are several methods for monitoring the position of 
A QPD is a typical example of a position detection system commonly used 
in optical tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is directed onto the QPD, with the maximum 
intensity mapping to the particle's centre of mass. By summing the 
horizontal and vertical quadrants together the particle's centre of mass 
is tracked in the x-y plane. While the tracking is accurate the QPD 
needs to be calibrated in order to convert the voltage signals into 
actual distance reading.

\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
Generation of fluid shear can be achieved via two avenues: Firstly, by 
utilising circularly polarised light it is possible to transfer angular 
momentum from the laser to the trapped entity. Secondly, one can 
directly move the trap within the imaging plane by steering the beam 
using either a galvanometric mirror or gimbal mirror. The following 
chapter outlines the work done with shear generated by circularly 
polarised light and the challenges of applying this to localising 
nucleation.

There are several options for particles that can be rotated using 
optical tweezers []. Over the course of this project two different micro 
spheres where investigated, vaterite and liquid crystal droplets. Both 
can be readily synthesised in the lab and are will rotate at a variety 
of sizes. While silver nano particles were considered their high cost 
and small size meant they were disregarded as an option for optical 
rotation.

\subsection{Rotation of Vaterite micro spheres}
Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping. 
Synthesis of vaterite micro spheres requires fine control of the nucleation process in order to maintain polymorphic control.

\section{Shear induced Nucleation}
It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e. propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event. Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off. This has been shown theoretically for both simple colloidal \textcolor{red}{[ , ]}  and ice crystal formation\textcolor{red}{[ ]}; however, no experimental work into these systems has been conducted to prove this is the case.  There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions. In \textcolor{red}{[ ]} they found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate. They tested the theoretical model established in \textcolor{red}{[ ]} which modifies the CNT to account for the effects of a nucleus undergoing shearing; they accounted for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus. There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised. However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius. In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing. This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.

Optical tweezing has often been used for micro-rheology, by computing the exact forces being exerted on the trapped sphere, one can determine the fluid viscosity of the local surrounding medium. Typically one would use a birefringent particle (i.e. vaterite) and rotating it within the fluid, the maximum rotation rate being a product of the fluid drag resisting the torque of the trapping beam. Likewise, one can use a galvanometric mirror to probe the drag force of the fluid, by understanding the trap strength (calibrating using a low frequency signal) one can compute the force required to escape the trap and likewise determine the drag force applied by a moving sphere. Likewise, one can perform the inverse experiment, by applying a known shear rate one can see how the fluid responds to the fluid shear; this is what we attempt with the galvano-mirror, how does a supersaturated solution respond to a localised fluid field? This requires that we both know the fluid velocity, and the shear rate applied by our probe particle. Furthermore, we also require an understanding of how the fluid is effected by the beam, both the temperature rise and the corresponding change to the fluid viscosity.

Understanding the fluid velocity around our trapped object is determined mostly by the Reynold's number of our system, for a sphere submersed in a moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}
Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's density and viscosity respectively. In our case we do not have a fluid moving around a sphere but a sphere moving through the fluid at some velocity $U$, assuming a no-slip boundary condition we can model the fluid velocity profile based on the velocity of the particle. There are two possible avenues for generating shear flow with a trapped particle; rotation of birefringent particles, and fluid flow induced by particle motion.

Rotating birefringent particles are by far the most common method for generating and measuring fluid flow in a solution. To see if we can even achieve the theoretical maximum shear rate, vaterite spheres were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm laser at set to 450 mW. The rotation frequency was determined using the QPD, and the particle sizes were computed by image analysis. With the particle size and rotation frequency, the tangential rotation speed is calculated via:

\begin{align}
	\label{eq:birefringent_speed}
	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
\end{align}

Where $d$ is the particle diameter, $\omega$ is the rotation frequency reported by the QPD, and $r$ is the distance from the particle's centre. Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating outward from the centre of the sphere. The shear rate can then be computed as the partial derivative fluid flow (assuming shearing is generated purely by the flow field):

\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}

\begin{figure}[h]
	\label{fig:vaterite_shear}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere. (b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line. Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW. Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
\end{figure}

From Fig.\ref{fig:vaterite_shear} it's clear that

Using a galvano mirror can A particle's motion can be precisely controlled For a simple circular path one can estimate the sphere's speed by the radius of its path and the frequency of its orbit $U = R\omega$; however for a more complex path, such as an elliptical orbit the curve needs to be parametrised. One can describe the position parameter of an ellipse as such:

\begin{align}
	r(u) = \left[acos(2\pi u),bsin(2\pi u), 0 \right]
\end{align}

where a and b are the different characteristic radii of an ellipse, if we assume that u describes time from some initial point we can say $u=t\omega$ where omega is the frequency of orbit. Subbing this in and then taking the partial derivative of position gives:

\begin{align}
	v(t) = \frac{dr(t)}{dt} = \left[-2\pi a\omega \ sin(2\pi t\omega),\ 2\pi b\omega \ cos(2\pi t\omega),0 \right]
\end{align}

In order to compute U we simply take the magnitude of our velocity. For low velocities the fluid flow around the entire sphere can be computed based on the sphere's velocity.

\begin{align}
	u_r(r)=-v(t)cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $\theta$ is the angle from the direction of movement to the point you wish to measure, and $r$ is the radial distance to that point. Again taking the partial derivative we can get the shear rate for a particle moving through the fluid:

\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = v(t)cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
\end{align}

Silica beads ($r=1,57 \mu m$) were trapped and moved along an elliptical path

>>>>

\setcounter{chapter}{2}
\include{main}

### 2024-06-07 Latex_files/Thesis/chapters/chapter 4/chapter4.tex
updating main and chapters using

\setcounter{chapter}{3}
\chapter{Complex Langevin dynamics of simple spherical aggregates}
Much of the calibration theory discussed in Chapter 2 assumes that the target particle in question is a single sphere, one who's scattering and motion is easily computed. However, while working with dense colloidal suspensions, one often ends up trapping more than one sphere. Li and Arlt \cite{Li2008} studied the case of two microspheres trapped in a single OT and found that multiple trapped beads could be mistaken for a single trapped bead with altered trap stiffness. Theoretical studies on the case of two trapped microspheres by Xu \textit{et.al.} \cite{Xu2005} employed a ray-optics based model to show that the two trapped beads are brought into physical contact with each other by optical forces and they also calculated the axial equilibrium positions of the two trapped beads as a function of their size. Experiments in \cite{Praveen2016} confirmed that the two trapped beads indeed experience different trap stiffnesses in the vicinity of the same potential well. There are further discussions looking into the dynamics of a whole host of asymmetrically shaped particles \cite{Loudet2014, ShengHua2005, Chetana2022}, their results all showing that predicting the behaviour an arbitrary shaped particle comes with great difficulty due to the fact that the optical force is dependent on a greater number of variables such as orientation and size factors.

In this chapter we consider how the addition of a second sphere into an optical trap can radically effect its dynamics, to the extent that one can no longer rely on typical calibration techniques to characterise the interactions.

\section{Positional and Orientational dependence of Trapping forces}
If we wanted to start from first principles and determine the trap strength on our target particle the first step would be to locate the harmonic traps relative to the trap focus. The methodology for computing optical forces has been covered extensively for a number of different trapping conditions \cite{RanhaNeves2019}, so it is relatively easy to compute the trapping force and determine where a simple sphere would be located relative to focal point of the laser. We can do so because the optical force is only dependent on the particle's relative position. If we instead consider a asymmetric dimer for example we see just by inverting the particle then a secondary harmonic trap can be found below the focus.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.475\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{.475\linewidth}
		\centering
		\raisebox{65pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\begin{subfigure}{.475\linewidth}
		\centering
		\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2. (a) is the case where the smaller sphere is orientated with the beam propagation direction. (b) is the inverted case, smaller sphere oriented against the propagation direction. Renders to visualise the dimer orientation are shown below each plot The black lines on each force-curve is a linear fit with the slope being reported as the trap stiffness in the legend.}
\end{figure}

We can see that the trap below the focus is comparable in strength to above the focus, however the difference in the transverse strength is far more noticeable. As shown below in Fig~\ref{fig:transverse_force}, the dimer's orientation and relative position significantly changes the force curve; not only is the trap wider when inverted but the trap stiffness is increased. This highlights one of the challenges involved with studying asymmetric particles, even though its a simple enough process to trap them they maybe characterised very differently depending on their relative position and orientation towards the trap. This can have a significant impact on rheological studies - or attempting to probe any local property - as the variance in trap strength can result in large errors over repeated measurements. 
\begin{figure}
	\centering
	\label{fig:transverse_force}
	\includegraphics[width=0.67\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the point of the contact of the spheres (µm) for the case of a dimer of size ratio 2 while being displaced in the transverse plane. With the blue curve representing the force response for a dimer in its standard orientation, orange being the inverted case, and green the same case but placed below the focus.}
\end{figure}

For completeness the harmonic traps were located for dimers across a range of size ratios - from $a_1/a_2 = 1$ to $a_1/a_2=10$ - while also recording the trap stiffness for each trap. As $a_2$ decrease the dimer begins to approximate a single homogenous sphere - at least in terms of location and trap strength. However, for intermediate sized dimers (between $a_1/a_2 = 1.1$ to $a_1/a_2=4$), a second harmonic trap appeared below the trapping focus. Previous work using the ray-optics model have confirmed even in the case that two spheres begin separated the electric field will align the molecules as such that they make contact and are trapped together about a single trapping position \cite{Xu2005}. Furthermore it has been shown through proper manipulation of the Gaussian or Bessel beam modes that any number of trapping potentials can be developed \cite{Shahabadi2020}. This result however, is the first example of an orientation dependent trapping situation using only a $TEM_00$ beam.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying size ratio, dotted lines represent unstable traps whereas solid lines represent stable trapping positions. (a) shows that dimers with their smaller sphere orientated away from the focus have an expected single trapping position. (b) shows that when the same dimer is inverted $180^\circ$ there are now stable traps along the beam axis, one below the focus and one above the focus.}
\end{figure}

This was only seen when the dimer is orientated with the smaller sphere above the larger sphere, when the dimer is flipped then the force curve only intersects the x-axis once, regardless of size ratio. More noteworthy is that in this orientation the dimer’s equilibrium position reaches its limit at a smaller size ratio; before it took a size ratio of 1:4 to achieve a final equilibrium position whereas now the second sphere needed to shrink until it was 1/7th of the first sphere’s size before the equilibrium position settled. This would suggest by having a second sphere behind the first, the momentum transfer from the beam to the dimer is diminished resulting in a weaker force pushing against the dimer as it sits within the trap. And when the dimer is flipped this shielding is lost and so both beads are being pushed by the beam.

Computing the equilibrium positions when a dimer is aligned with the electric field is relatively trivial as the orientational torque is minimised (see Eq.\ref{eq:opt_torque}), meaning once trapped the dimer is unlikely to change orientation enough to escape the trap. However, that does not rule out the possibility that there is a stable orientation that is not strictly vertical, in fact most experimental work with symmetric dimers will trap them lying perpendicular to the beam direction \cite{Ahn2018}. Unlike before where we can simply find the trap by varying the dimer's vertical position its instead more prudent to run a multitude of smaller simulations at a variety of starting positions and orientations. An example for a dimer of size ratio 2 is shown below:

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{off_axis_trap.png}
	\caption{Trajectory map of simulations ran using a dimer of size ratio 2 with a laser power of 500 mW. The stable points are indicated by the larger spheres and the starting conditions are colour coded to match the stable point they end up in.}
\end{figure}

Interestingly the trap strength of the these off-axis traps are similar in magnitude to the vertically aligned traps, but when the laser power is lowered (to ~5 mW) the traps become metastable resulting in the dimer escaping from after some random time in the trap. 
\bibliographystyle{ieeetran}
\bibliography{thesis_bib}

>>>>

\setcounter{chapter}{3}
\include{main}

\bibliographystyle{ieeetran}
\bibliography{thesis_bib}
